---
title: "Policy & Claims for Flood Insurance"
author: "Andrew Ba Tran"
date: "8/28/2017"
output: html_document
---

Harris County, which includes Houston, has received more than $3 billion in flood losses from the Federal Emergency Management Agency since 1976. It ranks third behind Orleans and Jefferson counties in Lousiana, which sustained significant damage during Hurricane Katrina. 

The county also ranks third for total losses with more than 100,000 total losses recorded— almost double the fourth-ranking Miami-Dade County.

As a city, Houston has received about $1.6 billion in flood mitigation assistance. The fourth-largest amount in the country.

From FEMA's National Flood Insurance Program [data page](https://www.fema.gov/policy-claim-statistics-flood-insurance). From 1978 to March 2017.

```{r setup, include=FALSE, warning=F, message=F, fig.width=10, fig.height=6}

library(tidyverse)
library(DT)
library(stringr)
# required functions
source("analyze_policies.R")  # function to preprocess policies data
source("analyze_claims.R")  # function to preprocess claims data

policies <- Fn_Analyze_Policies()
claims <- Fn_Analyze_Claims()

all_data <- merge(policies, claims, by = c("state", "county"), all = TRUE)
# convert state and county names to be consistent with ggplot2
all_data$state <- tolower(all_data$state)
all_data$county <- tolower(all_data$county)
# remove ' county' and ' parish' from county names
all_data$county <- gsub(" county", "", all_data$county)
all_data$county <- gsub(" parish", "", all_data$county)
```

# Losses

```{r table1, echo=F}
all_data_losses <- select(all_data, State=state, County=county, `Total Losses` = total_loss, `Closed losses` = closed_loss, `Open losses`=open_loss, `CWOP losses`=cwop_loss, `Total Payments`=total_pay) %>% 
  arrange(desc(`Total Payments`))
datatable(all_data_losses, extensions="Buttons", options=list(
  dom="Bfrtip",
  buttons=c("csv", "excel", "pdf")
))
```

# Policy stats

```{r table2, echo=F}
all_data_policies <- select(all_data, State=state, County=county, `Policies In-force`=policies, `Insurance In-force whole`=insurance, `Written Premium in-force`=premium) %>% 
  arrange(desc(`Written Premium in-force`))
datatable(all_data_policies, extensions="Buttons", options=list(
  dom="Bfrtip",
  buttons=c("csv", "excel", "pdf")))
```

# City summary

* Limiting to the 3,000 towns that received the most money *

```{r cities, echo=F, warning=F, message=F}
claims <- read.delim("data/formatted_claims_all.txt", as.is = TRUE)
claims <- filter(claims, !grepl("COUNTY*", community, perl=T))
claims$community <- gsub(", TOWN OF", "", claims$community)
claims$community <- gsub(", CITY OF", "", claims$community)

claims <- arrange(claims, desc(total_pay)) 
claims <- claims[1:3000,]

datatable(claims, extensions="Buttons", options=list(
  dom="Bfrtip",
  buttons=c("csv", "excel", "pdf")))

```

# Square miles

```{r table3, echo=F, warning=F, message=F}
all_data$state_county <- paste(all_data$state, all_data$county)
pop_land <- read_csv("data/DEC_10_SF1_GCTPH1.US05PR_with_ann.csv", skip=1)
pop_land$state <- gsub(".*United States - ", "", pop_land$`Geographic area`)
pop_land$state <- str_to_lower(gsub(" -.*", "", pop_land$state))

pop_land$county <- str_to_lower(gsub(" County", "", pop_land$`Geographic area_1`))
pop_land$state_county <- paste(pop_land$state, pop_land$county)

all_data <- left_join(all_data, pop_land)
colnames(all_data) <- make.names(colnames(all_data))

all_data$per_sqft_total <- round(all_data$total_pay / all_data$Area.in.square.miles...Total.area,2)
all_data$per_cap_total <- round(all_data$total_pay / as.numeric(all_data$Population),2)

all_data_perstuff <- select(all_data, state, county, `Per square mile` = per_sqft_total, `Per person` = per_cap_total)

datatable(all_data_perstuff, extensions="Buttons", options=list(
  dom="Bfrtip",
  buttons=c("csv", "excel", "pdf")
))
```

